Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2023-06-21 unimodular-covers-CombTheory-final.tex
typo
added comments to final correction to ref Y

\begin{claim}
\label{claim:r}
Assume without loss of generality that $b$ and $t$ either are parallel or their affine spans cross in $V_2$ (if they cross in $V_1$, the same claim can be reworded for $V_1$ and $l$). Then, 
\begin{enumerate}
\item The intersection of $Q$ with any line parallel to $p$ in $V_2$ has width with respect to $f_q$ strictly smaller than $w$.
\item $f_p(r_2) > f_p(r_1)$, that is, $f_p$ achieves its maximum over $r$ in $H_2$.
\end{enumerate}
\end{claim}

@@

We are now ready to show a contradiction. Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right. Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
 Let $r_1'$ and $r_2'$ be its endpoints.

We consider two cases, depending on whether $q_1+q_2-p_1$ lies ``to the right'' or ``to the left'' of $p_1$. That is, whether $f_p(q_1+q_2-p_1)$ is positive or negative. %For illustration use the left part of Figure~\ref{fig:claim4} in which $f_p(q_1+q_2-p_1)=0$, and in which both arguments work:

\begin{itemize}
\item \underline{$f_p(q_1+q_2-p_1)> 0$}. In this case the segment $r$ in the boundary of $P$ separates $p$ from  $q_1+q_2-p_1$.
Let $r''_1$ and $r''_2$ be the intersection points of the line containing $r$ with $b$ and $t$ respectively. We have that
$r'=[r'_1,r'_2]$ is parallel to $[r''_1,r''_2]$ but shorter, so that
\[
w= \width_{f_q}(p) \le \width_{f_q}(r) = \width_{f_q}(r') \le  f_q(r''_2) -f_q(r''_1), 
\]
where the first inequality comes from \eqref{eq:widthp}.
But on the other hand
\[
f_q(r''_2) < f_q(q_1+q_2-p_1) = 1, \quad f_q(r''_1) > f_q(p_1) =-1,
\]
which leads to the contradiction $w<2$.

\item \underline{$f_p(q_1+q_2-p_1)\ge 0$}.
Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ and $s_2$ on the line spanned by $t$. The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$, which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.

>>>>

\begin{claim}
\label{claim:r}
Assume without loss of generality that $b$ and $t$ either are parallel or their affine spans cross in $V_2$ (if they cross in $V_1$, the same claim can be reworded for $V_1$ and $l$). Then, 
\begin{enumerate}
\item The intersection of $Q$ with any line parallel to $p$ in $V_2$ has width with respect to $f_q$ strictly smaller than $w$.
\item $f_p(r_2) > f_p(r_1)$, that is, $f_p$ achieves its maximum over $r$ in $H_2$.
\end{enumerate}
\end{claim}
\Giulia{Do we ever use the claim 3(1), other than to prove 3(2)? If not, we can remove it from the statement...}

@@

We are now ready to show a contradiction. Since the normal fan of $Q$ refines that of $P$, the boundary of $Q$ must contain an edge parallel to $r$ on the right. Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
 Let $r_1'$ and $r_2'$ be its endpoints.

We consider two cases, depending on whether $q_1+q_2-p_1$ lies ``to the right'' or ``to the left'' of $p_1$. That is, whether $f_p(q_1+q_2-p_1)$ is positive or negative. %For illustration use the left part of Figure~\ref{fig:claim4} in which $f_p(q_1+q_2-p_1)=0$, and in which both arguments work:
\giulia{we put the origin in $q_1$, so the functional $f_p$ is not $0$ at $p$... so we could say $q_1 + q_2-p_1$ in $V_1$ or $V_2$... Though I'm not against changing our origin to lie in the intersection of $p$ and $q$}

\begin{itemize}
\item \underline{$f_p(q_1+q_2-p_1)> 0$}. In this case the segment $r$ in the boundary of $P$ separates $p$ from  $q_1+q_2-p_1$.
Let $r''_1$ and $r''_2$ be the intersection points of the line containing $r$ with $b$ and $t$ respectively. We have that
$r'=[r'_1,r'_2]$ is parallel to $[r''_1,r''_2]$ but shorter, \giulia{I would say that $r'$ needs to lie between the lines along $b$ and $t$ to explain why it's shorter} so that
\[
w= \width_{f_q}(p) \le \width_{f_q}(r) = \width_{f_q}(r') \le  f_q(r''_2) -f_q(r''_1), 
\]
where the first inequality comes from \eqref{eq:widthp}.
But on the other hand
\[
f_q(r''_2) < f_q(q_1+q_2-p_1) = 1, \quad f_q(r''_1) > f_q(p_1) =-1,
\]
which leads to the contradiction $w<2$.

\item \underline{$f_p(q_1+q_2-p_1)\le 0$}.
Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ and $s_2$ on the line spanned by $t$. The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$,\giulia{we still need to explain why this is true...} which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.
